---
title: "The Buffet Bet - Basket of Funds vs. All Passive"
subtitle: "Stats 107 Final Project"
author: "David Wihl"
date: "December 6, 2016"
abstract: |
  Ten years ago, Warren Buffett made a famous : the ultimate passive investment, the Vanguard S&P 500 Index Fund Admiral (VFIAX), vs. a basket of funds determined by a team of hedge fund experts. Whoever had a lower return after ten years would donate $1 million to the charity of choice of the winner. After eight years, it looks increasingly likely that Buffett will win the bet.
output: pdf_document
---
\begin{center}
\textbf{Type: Simulating a trading strategy.}
\end{center}

```{r, echo=FALSE, message=FALSE,warning=FALSE}
# Load libraries
library(ggplot2)
library(quantmod)
library(knitr)
```

# Introduction

Yahoo only has data on VFIAX starting 2000-11-13.

Quantmod does not use adjusted price to calculate returns. So, basically, it is
wrong. I found this going through the quantmod sources and then confirmed
it at http://stackoverflow.com/questions/34772616/r-function-periodreturn-not-computing-returns-using-adjusted-closing-prices

Must use OHLC numbers only



# Methods

## Importing List of Vanguard Mutual Funds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#
# Import List of Securities
#
loadRData = function(filename) {
  # Source: http://stackoverflow.com/questions/5577221/how-can-i-load-an-object-into-a-variable-name-that-i-specify-from-an-r-data-file
  load(filename)
  get(ls()[ls()!="filename"])
}
fromDate = "2010-11-13"
toDate   = "2016-11-13"
funds    = read.csv("data/vanguard.csv",header = T, stringsAsFactors = F)
# Load from previously cached download
stockDataEnv= loadRData("data/stockData.RData")
fundData = mget(funds$Ticker,stockDataEnv)
```

## Summary Statistics 

```{r, echo=FALSE}
#
# Generate Summary Statistics
#
for (i in 1:nrow(funds)) {
  sym = funds$Ticker[i]
  data = eval(parse(text=paste("fundData$",sym,sep="")))
  if(!is.null(data)){
    # we have the stockdata; Calculate summary statistics
    funds$startDate[i] = as.Date(index(data[1]))
    funds$endDate[i] = as.Date(index(last(data[])))
    startPrice = as.numeric(Ad(data[1]))
    endPrice = as.numeric(Ad(last(data[])))
    funds$totalRet[i] = (endPrice - startPrice) / startPrice
    dailyRets = dailyReturn(Ad(data))
    funds$stdDev[i] = sd(dailyRets)
    funds$avgRet[i] = mean(dailyRets)
    funds$Sharpe[i] = funds$avgRet[i] / funds$stdDev[i]
    # CAGR
    years =  as.numeric((as.Date(funds$endDate[i]) -
                           as.Date(funds$startDate[i])))/365
    funds$CAGR[i] = ((endPrice / startPrice)^(1/years)) - 1

  }
}
```

```{r, echo=FALSE}
#
# Generate Tables of Top Performers
#
colsToDisplay = c("Ticker","Fund.Name","Expenses","CAGR","Sharpe")
kable(head(funds[order(-funds$Sharpe),colsToDisplay]), caption="Best Sharpe Ratio",
      row.names = F,digits=2)

kable(head(funds[order(-funds$CAGR),colsToDisplay]), caption="Best CAGR",
      row.names = F,digits=2)

```

```{r, echo=FALSE}
#
# Show histograms with base VFIAX case
#
par(mfrow=c(3,1),pin=c(1.5,1.5))

# CAGR
hist(funds$CAGR,main="CAGR (all funds)",xlab="")
abline(v=funds[funds$Ticker=="VFIAX",]$CAGR,col="blue")

# Standard Deviation
hist(funds$stdDev,main="Standard Deviation (all funds)",xlab="")
abline(v=funds[funds$Ticker=="VFIAX",]$stdDev,col="blue")

# Sharpe Ratio
hist(funds$Sharpe,main="Sharpe Ratio (all funds)",xlab="")
abline(v=funds[funds$Ticker=="VFIAX",]$Sharpe,col="blue")
```


# Results

# Conclusions and Discussions

# References

This entire project code including source data can be found on GitHub ([https://github.com/wihl/stats107-project](https://github.com/wihl/stats107-project))



# Appendix A - Code

## Preparing Report

The following is the code used to prepare this report:
```{r code=readLines(knitr::purl('writeup.Rmd', documentation = 0)), eval = FALSE}

```

## Caching Stock Data

Stock data was downloaded and cached with the following code:
```{r, code=readLines("cache.R"), eval = FALSE}

```

