---
title: "The Buffet Bet - Basket of Funds vs. All Passive"
subtitle: "Stats 107 Final Project"
author: "David Wihl"
date: "December 6, 2016"
abstract: |
  \textbf{Type: Simulating a trading strategy.}
  Eight years ago, Warren Buffett made a famous bet: the ultimate passive investment, 
  the Vanguard S&P 500 Index Admiral Mutual Fund (VFIAX), vs. a basket of funds determined
  by a team of hedge fund experts. Whoever had a lower return after ten years 
  would donate $1 million to the charity of choice of the winner. After eight years, 
  it looks increasingly likely that Buffett will win the bet.
output: 
  pdf_document:
    toc: true
    number_sections: true
    fig_crop: false
linkcolor: blue
---

```{r, echo=FALSE, message=FALSE,warning=FALSE}
# Load libraries
library(ggplot2)
library(quantmod)
library(knitr)

# dot product
"%.%" <- function(x,y) sum(x*y)
# Calculate percent ready for display
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

# Constants
fromDate = "2000-11-13"
toDate   = "2016-11-13"
startCash = 100000
```

# Introduction

Can a basket of funds beat the market in the long term?

Our project aims to create a Mutual Fund selector for semi-passive investor. 
A typical Vanguard customer may have to choose from over one hundred mutual funds
without any guidance. We will create a portfolio to maximize returns over the 
long term using several common trading strategies, such as maximizing the
Sharpe ratio. In order to make this more realistic, we will take into account 
fees and transaction costs. Since this is aimed at a typical Vanguard retail customer, no shorting will be allowed. 

We will not account for taxes on the assumption that the funds are in a tax deferred account such as 401(k) or
IRA. For this reason, we will not evaluate ETFs which are at a disadvantage in a tax deferred account.

The portfolio will start with $100,000, which is the current average 401(k) balance 
(\href{http://time.com/money/4357248/retirement-401k-vanguard-balance/}{source}).

The overall time period will be evaluated whenever possible on a sixteen year history
which is the maximum provided by Yahoo via `quantmod`.

The baseline or benchmark will be investing this amount in VFIAX and leaving it 
there for the duration. The first section examines VFIAX to determine if it is a valid
proxy for market both in terms of similiar results and reasonable expenses. We will 
also examine how VFIAX compares to other Vanguard mutual funds in terms of CAGR and Volatility
in order to motivate several trading strategies.

The starting fund in the basket will be
VTHRX, which is one of the typical Vanguard Age-Adjusted balanced
funds and a common default fund for many retirement plans.

We will simulate whether rebalancing on a monthly, quarterly, semi-annual or annual basis is the best strategy.

Time permitting, we will evaluate these trading strategies over six windows of ten years each for a
more robust estimation of performance.

# Baseline

## Importing List of Vanguard Mutual Funds

A list was obtained of the 168 currently offered Vanguard mutual funds
including published expense ratios. The list was converted into a CSV for easy import into
R.

Once the list of funds was imported,  `quantmod` was used to read in
historical daily adjusted data for the entire date range available
from Yahoo. Since the earliest available date was November 13, 2000,
 an even 16 year end period was chosen, ending November 13, 2016. While more
data would have been preferable, it was felt that this was a reasonable sample
given it included two significant market corrections in 2001 and again in 2009.

Given the size of data (168 securities, 16 years of daily data), the data was cached
locally to allow fast iteration on many models. See appendix for all code.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#
# Import List of Securities
#
loadRData = function(filename) {
  # Source: http://stackoverflow.com/questions/5577221/how-can-i-load-an-object-into-a-variable-name-that-i-specify-from-an-r-data-file
  load(filename)
  get(ls()[ls()!="filename"])
}
funds    = read.csv("data/vanguard.csv",header = T, stringsAsFactors = F)
# Load from previously cached download
stockDataEnv= loadRData("data/stockData.RData")
fundData = mget(funds$Ticker,stockDataEnv)
```

## Summary Statistics 

Several simple descriptive statistics were created: Total Return, Average
Daily Return, Standard Deviation / Risk, Sharpe Ratio and CAGR. The following
tables summarize the best and worst mutual funds. 

Interestingly, in performing these summary statistics, it was determined that
`quantmod` does not use adjusted price to calculate returns. Without explicitly using adjusted pricing, the default
calculation is
wrong.  The was first discovered by going through the quantmod sources and then confirming at http://stackoverflow.com/questions/34772616/r-function-periodreturn-not-computing-returns-using-adjusted-closing-prices  As an extra precaution, values were adjusted upon retrieval using `getSymbol`.



```{r, echo=FALSE}
#
# Generate Summary Statistics
#
for (i in 1:nrow(funds)) {
  sym = funds$Ticker[i]
  data = eval(parse(text=paste("fundData$",sym,sep="")))
  if(!is.null(data)){
    # we have the stockdata; Calculate summary statistics
    funds$startDate[i] = as.Date(index(data[1]))
    funds$endDate[i] = as.Date(index(last(data[])))
    funds$startPrice[i] = as.numeric(Ad(data[1]))
    funds$endPrice[i] = as.numeric(Ad(last(data[])))
    funds$totalRet[i] = (funds$endPrice[i] - funds$startPrice[i]) / funds$startPrice[i]
    dailyRets = dailyReturn(Ad(data))
    funds$stdDev[i] = sd(dailyRets)
    funds$avgRet[i] = mean(dailyRets)
    funds$Sharpe[i] = funds$avgRet[i] / funds$stdDev[i]
    # CAGR
    years =  as.numeric((as.Date(funds$endDate[i]) -
                           as.Date(funds$startDate[i])))/365
    funds$CAGR[i] = ((funds$endPrice[i] / funds$startPrice[i])^(1/years)) - 1

  }
}
# Store some VFIAX data in separate variables for convenience
VFIAX.CAGR = funds[funds$Ticker=="VFIAX","CAGR"]
VFIAX.stdDev = funds[funds$Ticker=="VFIAX","stdDev"]
VFIAX.startPrice = funds[funds$Ticker=="VFIAX","startPrice"]
VFIAX.endPrice = funds[funds$Ticker=="VFIAX","endPrice"]
VFIAX.Sharpe = funds[funds$Ticker=="VFIAX","Sharpe"]
VFIAX.totReturn =funds[funds$Ticker=="VFIAX","totalRet"]

```

```{r, echo=FALSE}
#
# Generate Tables of Top Performers
#
colsToDisplay = c("Ticker","Fund.Name","Expenses","CAGR","Sharpe")
kable(head(funds[order(-funds$Sharpe),colsToDisplay],n=3), caption="Best Sharpe Ratio",
      row.names = F,digits=2)

kable(head(funds[order(funds$Sharpe),colsToDisplay],n=3), caption="Worst Sharpe Ratio",
      row.names = F,digits=2,n=3)

kable(head(funds[order(-funds$CAGR),colsToDisplay],n=3), caption="Best CAGR",
      row.names = F,digits=2,n=3)

kable(head(funds[order(funds$CAGR),colsToDisplay],n=3), caption="Worst CAGR",
      row.names = F,digits=2,n=3)

```

VMMSX is particularly bad: high expenses, low returns and low Sharpe Ratio.

## Evaluation of VFIAX

The following histograms show the distribution of all funds. VFIAX is the blue line.

```{r, echo=FALSE, fig.height=9.5}
#
# Show histograms with base VFIAX case
#
par(mfrow=c(3,1))

# CAGR
hist(funds$CAGR,main="CAGR (all funds)",xlab="",breaks=20)
abline(v=VFIAX.CAGR,col="blue")

# Standard Deviation
hist(funds$stdDev,main="Standard Deviation of Returns (all funds)",xlab="",breaks=20)
abline(v=VFIAX.stdDev,col="blue")

# Sharpe Ratio
hist(funds$Sharpe,main="Sharpe Ratio (all funds)",xlab="",breaks=20)
abline(v=VFIAX.Sharpe,col="blue")
```

Based on these summary statistics and distributions, it appears that there are better
mutual fund choices than VFIAX as a combination of higher CAGR and lower volatility.

## Expenses

Every mutual fund has expense ratios which is the principle way that a 
company like Vanguard makes revenue. An expense ratio is a mutual fund's 
annual operating expense, expressed as a percentage of the fund's average
net assets. It's calculated annually and removed from the fund's earnings
before they're distributed to investors, directly reducing investors'
returns. Since expenses were taken directly from assets, no special calculations
needed to be made for expenses, since the costs were already reflected in 
share prices, and therefore the returns.

For VFIAX, Vanguard claims a 0.05% expense ratio. These are
considered "admiral" class shares. For smaller investors, there
are "investor" class shares (VFINX) with a 0.16% expense ratio. 
This expense has gone
down over time and will likely continue to do so. 
We shall now examine if this matches reality
and how returns deviate from the benchmark S&P 500.

```{r, echo=FALSE, warning=FALSE,message=FALSE, results="hide"}
#
# Compare claimed Expense Ratios vs. Actual Expense Ratios
#
getSymbols("SPY", src="yahoo", from=as.Date(fromDate),to=toDate,adjust=T)
p0 = as.numeric(Ad(SPY[1]))
p1 = as.numeric(Ad(last(SPY[])))
tot = (p1 - p0)/p0

# Compound expenses over an expected 16 year period
expectedVFIAX = round((1.0005 ^ 16 - 1) * 100, 2)
expectedVFINX = round((1.0015 ^ 16 - 1) * 100, 2)
# Determine the percentage difference in total return 
spyOverVFIAX = round((tot - funds[funds$Ticker=="VFIAX",]$totalRet)/ funds[funds$Ticker=="VFIAX",]$totalRet*100,1)
spyOverVFINX = round((tot - funds[funds$Ticker=="VFINX",]$totalRet)/ funds[funds$Ticker=="VFINX",]$totalRet*100,1)
pctOverVFIAX = round((expectedVFIAX - spyOverVFIAX) / expectedVFIAX * 100,1)
pctOverVFINX = round((expectedVFINX - spyOverVFINX) / expectedVFINX * 100,1)

years =  as.numeric((as.Date(funds[funds$Ticker=="VFIAX",]$endDate)) -
                     as.Date(funds[funds$Ticker=="VFIAX",]$startDate  ))/365
spy.CAGR =  (p1 / p0)^(1/years)  - 1

```

Metric | VFINX (Investor) | VFIAX (Admiral)
-------|------------------|-----------------
Claimed Expense | 0.15% | 0.05%
Cum. Exp. 16 Yrs | `r round(expectedVFINX,1)`% |  `r round(expectedVFIAX,1)`% 
Actual Expense | `r spyOverVFINX`% | `r spyOverVFIAX`%
Difference | `r pctOverVFINX`% | `r pctOverVFIAX`%
CAGR | `r round(funds[funds$Ticker=="VFINX",]$CAGR*100,4)`% | `r round(funds[funds$Ticker=="VFIAX",]$CAGR*100,4)`%

SPY CAGR: `r round(spy.CAGR*100,4)`%


```{r, echo=FALSE}
# Plot Difference in Adjusted Price based on Expense Ratio
plot(Ad(SPY), main="S&P 500 vs. VFIAX, VFINX",ylab="Adjusted Price")
lines(Ad(fundData$VFINX),col="red",lwd=.1)
lines(Ad(fundData$VFIAX),col="blue",lwd=.5)
legend("bottomright",c("SPY","VFIAX","VFINX"),fill=c("black","blue","red"))
```

By these calculations, expenses add up over time but are lower than advertised. Admiral shares do not
offer significant benefit over Investor shares even though the expenses are purported to be 1/3. To be
precise, there are less than 2 basis points separating their CAGR.  For a very small cost, typically
1.5 basis points of CAGR for Admiral shares, Vanguard index funds offer significant convenience.


# Trading Methods

From the previous section, we have established that VFIAX:

* represents the S&P 500 market without significant overhead
* may not be an optimal choice. We should be able to do better.

We will now iterate through a number of trading strategies to determine which
trading models can best achieve this potential.

## Age Based Balanced Fund

The default mutual fund for many Vanguard-based retirement plans is age adjusted 
balanced fund, such as VTHRX. Since VTHRX is a newer fund, having started
on `r as.Date(funds[funds$Ticker=="VTHRX",]$startDate)`, we will calculate VFIAX
performance over the same time period.

```{r, echo=F}
#
# Compare VFIAX to VTHRX over equivalent time periods.
#
# Since VTHRX is younger than VFIAX, recalculate CAGR on matching dates
sDate = as.Date(funds[funds$Ticker=="VTHRX",]$startDate)
startPrice = as.numeric(Ad(fundData$VFIAX[sDate]))
endPrice = as.numeric(Ad(last(fundData$VFIAX[])))
totRet = (endPrice - startPrice) / startPrice
years = as.numeric((as.Date(funds[funds$Ticker=="VTHRX",]$endDate) -
                           as.Date(funds[funds$Ticker=="VTHRX",]$startDate)))/365
CAGR = ((endPrice / startPrice)^(1/years)) - 1
f1 = as.numeric(Ad(fundData$VFIAX[paste0(sDate,"::"),])) / startPrice
f2 = as.numeric(Ad(fundData$VTHRX)) / as.numeric(Ad(fundData$VTHRX[1]))
```

Metric | VFIAX  | VTHRX
-------|------------------|-----------------
Expense Ratio |  `r percent(funds[funds$Ticker=="VFIAX",]$Expenses/100)` |  `r percent(funds[funds$Ticker=="VTHRX",]$Expenses/100)`
CAGR | `r percent(CAGR)` | `r percent(funds[funds$Ticker=="VTHRX",]$CAGR)` 

Not looking so good: lower CAGR and higher expenses.

```{r, echo=FALSE}
plot(index(fundData$VTHRX),f1,type="l",main="VTHRX vs. VFIAX (Normalized)",xlab="",ylab="Adjusted Price",col="red")
lines(index(fundData$VTHRX),f2,col="blue",lwd=1)
abline(h=1)
legend("bottomright",c("VFIAX","VTHRX"),fill=c("red","blue"))
```

The age adjusted balanced fund slightly outperforms the market during downturns, but over a longer time period significantly
underperforms the market.

## Better Single Fund

From the selection of funds, is there a fund that outperforms the market by 20%, has been in existance
at least 10 years and has lower volatility?

```{r, echo=F}
#
# Find a better single fund that beats the market with higher CAGR and lower volatility
#
betterFunds = c("VFIAX")
for (i in 1:nrow(funds)) {
  if (funds$CAGR[i] > (1.2 * VFIAX.CAGR)) {
    if (funds$stdDev[i] < VFIAX.stdDev) {
      if (funds$startDate[i] < as.Date("2006-01-01")) {
        #cat(funds$Ticker[i]," ",funds$Fund.Name[i],"\n")
        betterFunds = c(betterFunds, funds$Ticker[i])
      }
    }
  }
}

colsToDisplay = c("Ticker","Fund.Name","startDate","Expenses","CAGR","stdDev")
bf = funds[funds$Ticker %in% betterFunds,colsToDisplay]
bf$startDate = as.Date(bf$startDate)
```

It turns out that there are `r nrow(bf)-1` out of `r nrow(funds)` that meet these criteria. Here are the top ten:

```{r, echo=F}
kable(rbind(head(bf[order(-bf$CAGR),],n=10), tail(bf[order(-bf$CAGR),],n=1)),
        caption="Top 10 Overall better funds than VFIAX",
      row.names = F,digits=3)

```

Let's plot a selection:

```{r, echo=F}
f1 = as.numeric(Ad(fundData$VFIAX)) / as.numeric(Ad(fundData$VFIAX[1]))
f2 = as.numeric(Ad(fundData$VASVX)) / as.numeric(Ad(fundData$VASVX[1]))
f3 = as.numeric(Ad(fundData$VGHCX)) / as.numeric(Ad(fundData$VGHCX[1]))
f4 = as.numeric(Ad(fundData$VWESX)) / as.numeric(Ad(fundData$VWESX[1]))
f5 = as.numeric(Ad(fundData$VHGEX)) / as.numeric(Ad(fundData$VHGEX[1]))

plot(index(fundData$VFIAX),f2,type="l",main="VFIAX vs. Better Funds (Normalized)",xlab="",ylab="Adjusted Price",col="blue")
lines(index(fundData$VFIAX),f1,col="black",lwd=0.5)
lines(index(fundData$VFIAX),f3,col="red",lwd=0.5)
lines(index(fundData$VFIAX),f4,col="green",lwd=0.5)
lines(index(fundData$VFIAX),f5,col="orange",lwd=0.5)
abline(h=1)
legend("topleft",c("VFIAX","VASVX","VGHCX","VWESX","VHGEX"),fill=c("black","blue","red","green","orange"))
```

We've discovered that approximately 15% of funds consistently beat the market over a 16 year period across
a large number of different asset classes: Health Care Sector, Consumer Staples Sector, even several bond funds.

Should the entire portfolio be put into one of these such as VASVX? Even though the volatility is lower, this would not be wise. Let's
now examine more complex strategies that involve diversification to see if risks can be lowered even further at while
potentially increasing in CAGR.

## Diversified Static Set of Funds

Vanguard and many other consumer general advice experts recommend a diversified set of static funds. Using
Vanguard supplied tools on their website, the following simulation was run using a mid-career individual
saving for retirement:

* Time to invest before withdrawl: $\geq$ 15 years
* Money will be withdrawn: $\geq$ 15 years
* Medium risk tolerance
* Medium likelihood of selling riskier assets during market declines
* Medium volatility
* Current and future income sources will be somewhat stable (3 on a scale of 5)
* $100,000 to invest

The Vanguard calculator came back with the following recommendations:

Weight | Ticker
-------|-------
36% | VTSMX
24% | VGTSX
28% | VBMFX
12% | VTIBX

Unfortunately VTIBX, the Total International Bond Index Fund Investor shares has only been available since
June 2013. Using Morningstar, an equivalent was found that had a sixteen year history: GMHBX, 
GMO Currency Hedged International Bond Fund Class III (GMO funds are unlikely to be available to a small
investor so this is a best case scenario for Vanguard suggestions.)

```{r, echo=FALSE,results="hide"}
#
# Compare against a diversified set of funds recommended by Vanguard
#
suggestions = c("VTSMX","VGTSX","VBMFX","GMHBX")
getSymbols("GMHBX", src="yahoo", from=as.Date(fromDate),to=toDate,adjust=T)
weights = c(.36,.24,.28,.12)
# Build covariance matrix
returns = cbind(dailyReturn(fundData$VTSMX),dailyReturn(fundData$VGTSX),dailyReturn(fundData$VBMFX),dailyReturn(GMHBX))
# delete one extra row from VBMFX
returns = returns[index(returns) !='2001-11-22']
fundData$VBMFX = fundData$VBMFX[index(fundData$VBMFX) != '2001-11-22']
names(returns) = suggestions
cov.mat = cov(returns)  # annualized
risk.p = sqrt(t(weights) %*% cov.mat %*% weights)
```

The portfolio risk is `r round(risk.p,4)` which is certainly better than VFIAX's `r round(VFIAX.stdDev,4)`

```{r, echo=FALSE}
shares = rep(0,length(suggestions))
startValue = rep(0,length(suggestions))
endValue = rep(0,length(suggestions))
CAGR = rep(0,length(suggestions))
# Find total return on three known funds
for (i in 1:length(suggestions)-1) {
  startValue[i] = startCash * weights[i]
  shares[i] = startValue[i] / funds[funds$Ticker==suggestions[i],"startPrice"]
  endValue[i] = shares[i] * funds[funds$Ticker==suggestions[i],"endPrice"]
}
# special case for GMHBX
startValue[4] = startCash * weights[4]
shares[4] = startCash * weights[4] / as.numeric(Ad(first(GMHBX)))
endValue[4] = shares[4] * Ad(last(GMHBX))
years =  as.numeric((as.Date(funds[funds$Ticker=="VFIAX","endDate"]) -
                           as.Date(funds[funds$Ticker=="VFIAX","startDate"])))/365
for (i in 1:length(suggestions)) {
  CAGR[i] = 100*(((endValue[i] / startValue[i])^(1/years)) - 1)
}
port = cbind(startValue, endValue,CAGR)
port.CAGR = 100*((sum(endValue) / startCash)^(1/years) - 1)
port = rbind(port, c(sum(startValue), sum(endValue), port.CAGR ))
port = rbind(port, c(startCash,startCash * (1+VFIAX.totReturn), VFIAX.CAGR*100))
rownames(port) = c(suggestions,"Total Portfolio","VFIAX")
colnames(port) = c("Start Value","End Value","CAGR")
port.prices = cbind(as.numeric(Ad(fundData$VTSMX)), as.numeric(Ad(fundData$VGTSX)),
                    as.numeric(Ad(fundData$VBMFX)), as.numeric(Ad(GMHBX)))

port.value = sweep(port.prices, MARGIN=2, shares,'*')
kable(port, caption="Vanguard Recommended Portfolio Growth", digits=2)

```

The recommended portfolio provide effectively the same returns as the market, albeit at
lower volatility. Note from the plot below that the portfolio had a better 
return than the market for the majority
of the period.

```{r, echo=FALSE}
f1 = as.numeric(Ad(fundData$VFIAX)) / as.numeric(Ad(fundData$VFIAX[1]))
f2 = port.prices[,1] / port.prices[1,1]
f3 = port.prices[,2] / port.prices[1,2]
f4 = port.prices[,3] / port.prices[1,3]
f5 = port.prices[,4] / port.prices[1,4]
f6 = rowSums(port.value) / startCash

plot(index(fundData$VFIAX),f2,type="l",main="VFIAX vs. Recommend Portfolio (Normalized)",xlab="",ylab="Adjusted Price",col="blue")
lines(index(fundData$VFIAX),f1,col="black",lwd=0.5)
lines(index(fundData$VFIAX),f3,col="purple",lwd=0.5)
lines(index(fundData$VFIAX),f4,col="green",lwd=0.5)
lines(index(fundData$VFIAX),f5,col="orange",lwd=0.5)
lines(index(fundData$VFIAX),f6,col="red",lwd=2.0)
abline(h=1)
legend("topleft",c("VFIAX","VTSMX","VGTSX","VBMFX","GMHBX","Portfolio"),fill=c("black","blue","purple","green","orange","red"))
```

## Minimum Variance

The original Buffet Bet was the market vs. a hedge fund consisting of a basket of funds. In retrospect,
it is not surprising that the market is beating the hedge fund. The primary purpose of the hedge fund
is the eponymous hedging - that is reducing risk. Hedge funds outperform the market during downtimes
but underperform over the long term. Their primary raison d'être is captial preservation. This strategy
will be mimicked by selecting a minimum variance portfolio. The portfolio will be compared against 
market downturn periods vs. long term.

```{r, echo=FALSE}
source('zivot_code.R')
#
# Minimum Variance portfolio
#
# We are now going to use the Zivot code to also build the efficient frontier
tickers = funds$Ticker
# Build covariance matrix

# TODO. Should we use Zivot? Is there a better way? How can we find optimal weights?
#returns = cbind(dailyReturn(fundData$VTSMX),dailyReturn(fundData$VGTSX),dailyReturn(fundData$VBMFX),dailyReturn(GMHBX))
#names(returns) = tickers
#cov.mat = cov(returns)  
#risk.p = sqrt(t(weights) %*% cov.mat %*% weights)


```

## CAPM / Sharpe Ratio / Efficient Frontier

As highlighted earlier in this paper, there are funds such as VSIAX which offer CAGR three times the market
albeit at higher risk (See Table 3). By using CAPM, a selection of funds will be made to maximize returns while minimizing risk.

<!--


## Pairs Trading

## Moving Averages

## Momentum Trading

## Fama French Three Factor

## Fama French Five Factor

-->

# Future Ideas

This paper has illustrated several techniques to beating the market over the long term using
well known trading techniques. It would be interesting to explore more complex novel techniques
that apply modern Machine Learning methods. One area particularly worth exploring would be
Reinforcement Learning which is well suited to problems such as maximizing returns while minimizing risks. 
Furthermore, additional factors like restrictions on number of trades,
and cost of trades are relatively easy to model.

# Conclusions and Discussions

In summary: VFIAX vs:

Alternative | Result | Comment
------------|--------|--------
Age Balanced | Loser | Lower CAGR, higher expenses
Single Alternative Fund | Winner | Higher return, but possibly higher risk
Diversified Static | Winner | Same return, significantly lower risk

This illustrates that there are a number of different techniques of besting the market
over the long term, either in terms of higher returns, lower risk or both.

# References

This entire project code including source data can be found on GitHub ([https://github.com/wihl/stats107-project](https://github.com/wihl/stats107-project))

Ang, Clifford S., _Analyzing Financial Data and Implementing Financial Models Using R_, Springer 2015,
ISBN 978-3-319-14075-9 (eBook)

List of Vanguard funds, including fees and transaction costs, supplied by Vanguard https://investor.vanguard.com/mutual-funds/list#/mutual-funds/asset-class/month-end-returns

The Buffet Bet, http://longbets.org/362/

Fortune, The Buffett Bet http://fortune.com/2016/05/11/warren-buffett-hedge-fund-bet/


# Appendix A - Code

## Preparing Report

The following is the code used to prepare this report:
```{r code=readLines(knitr::purl('writeup.Rmd', documentation = 0)), eval = FALSE}

```

## Caching Stock Data

Stock data was downloaded and cached with the following code:
```{r, code=readLines("cache.R"), eval = FALSE}

```

